---
title: "Project1"
author: "Ye Chao"
date: "2019/10/15"
output:
  html_document: default
  pdf_document: default
---
## Import data ----
```{r}
## Import ---- 
Trip <- read.csv("C:/Users/chaoye/Desktop/Davis/quarter 4/Graduate Project/2017 NHTS Data/trippub.csv", header = TRUE)
```

## I. Descriptive Analysis ----
### 0. Data pre-precessing ----
```{r}
# ### interested variables ----
# Trip$CDIVMSAR  #Grouping of household by combination of census division, MSA status, and presence of a subway system when population greater than 1 million
# Trip$HHSTATE  # state
# Trip$HHSTFIPS  # state fips
# 
# Trip$MSACAT  # Metropolitan Statistical Area (MSA) category for the household's home address, based on household's home
# Trip$MSASIZE # Population size category of the Metropolitan Statistical Area (MSA), from the 2010-2014 five-year American Community Survey (ACS) API.
# 
# 
# Trip$DRIVER  # Driver status
# Trip$DRVRCNT  #Number of drivers in household
# 
# Trip$HBPPOPDN # Category of population density (persons per square mile) in the census block group of the household's home location
# 
# Trip$LOOP_TRIP # Trip od same destination
# 
# Trip$NUMTRANS  # count of transfers
# Trip$TRACCTM # trip time to transit station in minutes
# Trip$TRWAITTM # transit wait time in minutes

## Variables selection ----
Myvars <- c("HOUSEID", "PERSONID", "TDTRPNUM", "TRPTRANS", "WHYTRP1S", "STRTTIME", "ENDTIME", "HBHUR", "TRVLCMIN", "TRPMILES", "GASPRICE", "HHFAMINC", "HHVEHCNT", "HHSIZE", "R_AGE", "R_SEX", "CDIVMSAR", "HHSTATE", "HHSTFIPS", "MSACAT", "MSASIZE", "DRIVER", "DRVRCNT", "HBPPOPDN", "LOOP_TRIP", "NUMTRANS", "TRACCTM", "TRWAITTM")
Mytrip <- Trip[Myvars]



## Mode selection ----
walk <- subset(Mytrip, Mytrip$TRPTRANS == 1)
bicy <- subset(Mytrip, Mytrip$TRPTRANS == 2)
pt <- subset(Mytrip, Mytrip$TRPTRANS == 11 | Mytrip$TRPTRANS ==12)
taxi <- subset(Mytrip, Mytrip$TRPTRANS == 17)


Mode <- rbind(walk, bicy, pt, taxi)

Mode <- subset(Mode, Mode$TRVLCMIN >= 0)
Mode <- subset(Mode, Mode$TRPMILES >= 0)
Mode <- subset(Mode, Mode$HHFAMINC >= 0)
Mode <- subset(Mode, Mode$R_SEX >= 0)

Mode$R_SEX <- as.factor(Mode$R_SEX)
levels(Mode$R_SEX) <- c("male", "female") # 0 as male, 1 as female.


# urban work trip ----
Mode.u.w <- subset(Mode, as.character(Mode$HBHUR) == "U" & Mode$WHYTRP1S == 10)
table(Mode.u.w$TRPTRANS)

walk <- subset(Mode.u.w, Mode.u.w$TRPTRANS == 1)
walk$TDTRPNUM <- 1:1702

bicy <- subset(Mode.u.w, Mode.u.w$TRPTRANS == 2)
bicy$TDTRPNUM <- 1:276

pt <- subset(Mode.u.w, Mode.u.w$TRPTRANS == 11 | Mode.u.w$TRPTRANS ==12)
pt$TDTRPNUM <- 1:461

taxi <- subset(Mode.u.w, Mode.u.w$TRPTRANS == 17)
taxi$TDTRPNUM <- 1:122

car <- subset(Mode.u.w, Mode.u.w$TRPTRANS == 3| Mode.u.w$TRPTRANS == 4|Mode.u.w$TRPTRANS == 5|Mode.u.w$TRPTRANS == 6)

Mode.u.w <- rbind(walk, bicy, pt, taxi)

Mode.u.w$TRPTRANS <- as.factor(Mode.u.w$TRPTRANS)
levels(Mode.u.w$TRPTRANS) <- c("walk", "bicy", "pt", "pt", "taxi")


## By the way, how many trips by private cars ----
car <- subset(Mytrip, Mytrip$TRPTRANS == 3 | Mytrip$TRPTRANS == 4 | Mytrip$TRPTRANS == 5 | Mytrip$TRPTRANS == 6)
car <- subset(car, as.character(car$HBHUR) == "U" & car$WHYTRP1S == 10)
nrow(car)
```

### 1. about the sample ----
```{r}
## MSA area
msa <- table(Mode.u.w$HHSTATE, Mode.u.w$MSASIZE)
msa <- as.data.frame.matrix(msa)
(sum(msa$`4`)+sum(msa$`5`))/sum(msa) # = 96.33% of the sample is from a MSA that has a population size more than 1 million. 

## Loop Trip or not
loop <- table(Mode.u.w$LOOP_TRIP)
loop <- as.data.frame(loop)
loop[2,]/sum(loop$Freq) # = 95.67% of the trips took place within the project scope are not loop trip since the purpose for the trips is going to work. 


```

### 2. when will they travel ----
```{r, , fig.width=8, fig.height=10}
# Mode.u.w$STRTTIME <- formatC(Mode.u.w$STRTTIME, width = 4, format = "d", flag = "0")
# Mode.u.w$ENDTIME <- formatC(Mode.u.w$ENDTIME, width = 4, format = "d", flag = "0")
# 
# Mode.u.w$STRTTIME<- format(strptime(Mode.u.w$STRTTIME, format="%H%M"), format = "%H:%M")
# Mode.u.w$ENDTIME<- format(strptime(Mode.u.w$ENDTIME, format="%H%M"), format = "%H:%M")
# 
# Mode.u.w$STRTTIME <- as.POSIXct(Mode.u.w$STRTTIME,format= "%H:%M")
# Mode.u.w$ENDTIME <- as.POSIXct(Mode.u.w$ENDTIME,format= "%H:%M")
# 
# # Mode.u.w$STRTTIME <- format(Mode.u.w$STRTTIME, "%H:%M:%S")
# # Mode.u.w$ENDTIME <- format(Mode.u.w$ENDTIME, "%H:%M:%S")

# install.packages("timelineS")
library(timelineS)
library(ggplot2)

timelineG(df=Mode.u.w, start="STRTTIME", end="ENDTIME", names="TDTRPNUM", group1="TRPTRANS", group2="R_SEX", )
```

### 3. They don't drive because they don't have driver ----
```{r}
library(ggplot2)
table(Mode.u.w$DRIVER)
table(Mode.u.w$HHVEHCNT)

driver <- subset(Mode.u.w, Mode.u.w$DRIVER > 0)
driver$DRIVER <- as.factor(driver$DRIVER)
levels(driver$DRIVER) <- c("Have a license", "Don't have a license")
driver$HHVEHCNT <- as.numeric(driver$HHVEHCNT)

## 3.1 with respect to family income ----
ggplot(data = driver) +
  geom_boxplot(mapping = aes(TRPTRANS, HHFAMINC))+
  labs(x= "Mode",y = "Household family income")+
  labs(title = "Bar Chart", subtitle = "Household family income - Mode")


# ggplot(data = driver, mapping = aes(x = TDTRPNUM, y = HHVEHCNT)) +
#    geom_boxplot(mapping = aes(color = TRPTRANS))+
#    facet_wrap( ~TRPTRANS)

## 3.2 vehicle count ----
veh <- aggregate(driver$HHVEHCNT~driver$TRPTRANS, FUN=mean)
veh <- as.data.frame.matrix(veh)


ggplot(data = veh, mapping = aes(x = veh$`driver$TRPTRANS`, y = veh$`driver$HHVEHCNT`)) +
  geom_col(mapping = aes(fill = veh$`driver$TRPTRANS`))+
  labs(x= "Mode",y = "Average household vehicle counts")+
  labs(title = "Average household vehicle counts - Mode")+
  labs(fill = "Mode")


# table(driver$DRVRCNT)

## 3.3 over household size ----
r1 <- aggregate(driver$DRVRCNT/driver$HHSIZE, by = list(driver$TRPTRANS), mean)
r2 <- aggregate(driver$HHVEHCNT/driver$HHSIZE, by = list(driver$TRPTRANS), mean)
r <- cbind(r1,r2)
r <- r[,-3]
names(r) <- c("Mode","Average drivercount / HHsize","Average vehiclecount / HHsize")
df2 <- reshape::melt(r, id = c(1))
ggplot(df2, aes(x=Mode, y=value, fill=variable)) +
    geom_bar(stat='identity', position='dodge')+
  labs(x= "Mode",y = "Ratio value")+
  labs(title = "Average driver count & vehicle count / Household size")
# r1 <- as.data.frame(table(driver$DRVRCNT/driver$HHSIZE))
# r1
# 
# ggplot(data = r) +
#   geom_col(mapping = aes(x = `1`, y = r$`2`)) +
#   geom_col(mapping = aes(x = `1`, y = r$`4`))
# 
# ggplot(r, aes(x=`1`, y=value, fill=variable)) +
#     geom_bar(stat='identity', position='dodge')
# 
# r2 <- as.data.frame(table(driver$HHVEHCNT/driver$HHSIZE))
```

### 4. About transfering and waiting ----
```{r}
# Trip$NUMTRANS  # count of transfers
# Trip$TRACCTM # trip time to transit station in minutes
# Trip$TRWAITTM # transit wait time in minutes
## 4.1 Number of transfer
Mode.u.w$NUMTRANS <- as.numeric(Mode.u.w$NUMTRANS)
table(Mode.u.w$NUMTRANS)
table(pt$NUMTRANS)

trans <- subset(pt, pt$NUMTRANS >= 0)
trans$TDTRPNUM <- 1:nrow(trans)
t1 <- table(trans$NUMTRANS)
t1 <- as.data.frame(t1)


ggplot(t1, mapping = aes(x=Var1, y=Freq)) +
  geom_bar(stat = "identity")+
  labs(x= "Number of transfer",y = "Frequency")+
  labs(title = "For the people taking public transit", subtitle = "Number of their transfer")



# table(walk$TRWAITTM)
# table(bicy$TRWAITTM)
# table(taxi$TRWAITTM)
# table(pt$TRWAITTM)
```

### 5. percetage stack plot: Mode - Income ----
```{r}
# mode & income
Myvars <- c("TRPTRANS", "HHFAMINC")

stackplot1 <- Mode.u.w[Myvars]
stackplot1$type <- NA

stackplot1$HHFAMINC <- as.factor(stackplot1$HHFAMINC)
levels(stackplot1$HHFAMINC) <- c("14999 or less", "14999 or less", "15000 - 34999", "15000 - 34999", "35000 - 74999", "35000 - 74999", "75000 - 124999", "75000 - 124999" , "more than 125000", "more than 125000", "more than 125000")

s1 <- as.data.frame.matrix(table(stackplot1$HHFAMINC, stackplot1$TRPTRANS))
s1 <- cbind(rownames(s1), data.frame(s1, row.names=NULL))
s2 <- reshape::melt(s1, id = c(1))
names(s2) <- c("Income level", "variable", "value")

ggplot(s2, aes(x = variable, y = value, fill = `Income level`)) + 
  geom_bar(position = "fill",stat = "identity") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x= "Mode",y = " ")+
  labs(title = "Percent of income categories within different modes")


```

### 6. percetage stack plot: Mode - age ----
```{r}
Myvars <- c("TRPTRANS", "R_AGE")
stackplot2 <- Mode.u.w[Myvars]
stackplot2 <- subset(stackplot2, stackplot2$R_AGE > 0)
stackplot2$type <- 0

for (i in 1:nrow(stackplot2)) {
  if (stackplot2$R_AGE[i] <= 18) {
  stackplot2$type[i] <- 1
  }
}

for (i in 1:nrow(stackplot2)) {
  if (stackplot2$R_AGE[i] > 18 & stackplot2$R_AGE[i] <= 30) {
  stackplot2$type[i] <- 2
  }
}

for (i in 1:nrow(stackplot2)) {
  if (stackplot2$R_AGE[i] > 30 & stackplot2$R_AGE[i] <= 40) {
  stackplot2$type[i] <- 3
  }
}

for (i in 1:nrow(stackplot2)) {
  if (stackplot2$R_AGE[i] > 40 & stackplot2$R_AGE[i] <= 50) {
  stackplot2$type[i] <- 4
  }
}

for (i in 1:nrow(stackplot2)) {
  if (stackplot2$R_AGE[i] > 50 & stackplot2$R_AGE[i] <= 60) {
  stackplot2$type[i] <- 5
  }
}

for (i in 1:nrow(stackplot2)) {
  if (stackplot2$R_AGE[i] > 60) {
  stackplot2$type[i] <- 6
  }
}

stackplot2$type <- as.factor(stackplot2$type)
levels(stackplot2$type) <- c("0-18", "18 - 30", "30 - 40", "40 - 50", "50 - 60", "Above 60")
# levels(stackplot2$type)

a1 <- as.data.frame.matrix(table(stackplot2$type, stackplot2$TRPTRANS))
a1 <- cbind(rownames(a1), data.frame(a1, row.names=NULL))
a2 <- reshape::melt(a1, id = c(1))
names(a2) <- c("Age", "variable", "value")

ggplot(a2, aes(x = variable, y = value, fill = Age)) + 
  geom_bar(position = "fill",stat = "identity") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x= "Mode",y = " ")+
  labs(title = "Percent of age categories within different modes")

```
### 7. Income ----
```{r}
## Income - vehicle ---- 

Myvars <- c("HHFAMINC", "HHVEHCNT")

stackplot3 <- Mode.u.w[Myvars]
# stackplot3$type <- NA

stackplot3$HHFAMINC <- as.factor(stackplot3$HHFAMINC)
levels(stackplot3$HHFAMINC) <- c("14999 or less", "14999 or less", "15000 - 34999", "15000 - 34999", "35000 - 74999", "35000 - 74999", "75000 - 124999", "75000 - 124999" , "more than 125000", "more than 125000", "more than 125000")

for (i in 1:nrow(stackplot3)) {
  if (stackplot3$HHVEHCNT[i] >= 3) {
    stackplot3$HHVEHCNT[i] <- 3
  }
}

table(stackplot3$HHVEHCNT)

v1 <- as.data.frame.matrix(table(stackplot3$HHFAMINC, stackplot3$HHVEHCNT))
v1 <- cbind(rownames(v1), data.frame(v1, row.names=NULL))
names(v1) <- c("Income", "0", "1", "2", "3 or more")


v2 <- reshape::melt(v1, id = c(1))
names(v2) <- c("Income level", "variable", "value")

ggplot(v2, aes(x = variable, y = value, fill = `Income level`)) + 
  geom_bar(position = "fill",stat = "identity") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x= "Car ownership",y = " ")+
  labs(title = "Percent of income categories within car ownership")
```
### 8. On and off peak ----
```{r}
## On and off peak ---- :

Mode.u.w$onoffid <- 0   #### peak hour idtificaiton: peak hour 1, off-peak hour 0.  

for (i in 1:nrow(Mode.u.w)) {
  if (Mode.u.w$STRTTIME[i] >= 600 & Mode.u.w$STRTTIME[i] <=900) {
    Mode.u.w$onoffid[i] <- 1
  }
 
  if (Mode.u.w$STRTTIME[i] >= 1700 & Mode.u.w$STRTTIME[i] <= 1900) {
    Mode.u.w$onoffid[i] <- 1
  }
}

onoff <- as.data.frame.matrix(table(Mode.u.w$TRPTRANS, Mode.u.w$onoffid))
onoff <- cbind(rownames(onoff), data.frame(onoff, row.names=NULL))
names(onoff) <- c("Mode", "Off-peak hour trip", "Peak hour trip")

onoff2 <- reshape::melt(onoff, id = c(1))
names(onoff2) <- c("Mode", "Peak hour", "Trip counts")


ggplot(onoff2, aes(fill=`Peak hour`, y=`Trip counts`, x=Mode)) + 
  geom_bar(position="dodge", stat="identity") +
  labs(title = "On and off peak hour trips for each mode")
  
```

## Descriptive ends ----
```{r}
rm(list = ls()) # empty the environment
```



















## II. Analytical Begins ---- 
### 0. Data pre-processing ----
```{r, echo=FALSE}
## Import ---- 
Trip <- read.csv("C:/Users/chaoye/Desktop/Davis/quarter 4/Graduate Project/2017 NHTS Data/trippub.csv", header = TRUE)
```


```{r}

# ### interested variables ----


# 1. Trip$MSACAT  # Metropolitan Statistical Area (MSA) category for the household's home address, based on household's home

# 2. Trip$HBPPOPDN #population density of household


##Variables selection ----
Myvars <- c("HOUSEID", "PERSONID", "TDTRPNUM", "TRPTRANS", "WHYTRP1S", "STRTTIME", "ENDTIME", "HBHUR", "TRVLCMIN", "TRPMILES", "GASPRICE", "HHFAMINC", "HHVEHCNT", "HHSIZE", "R_AGE", "R_SEX", "MSACAT", "HBPPOPDN")
Mytrip <- Trip[Myvars]


## Mode selection ----
walk <- subset(Mytrip, Mytrip$TRPTRANS == 1)
bicy <- subset(Mytrip, Mytrip$TRPTRANS == 2)
pt <- subset(Mytrip, Mytrip$TRPTRANS == 11 | Trip$TRPTRANS ==12)
taxi <- subset(Mytrip, Mytrip$TRPTRANS == 17)



Mode <- rbind(walk, bicy, pt, taxi)
Mode <- subset(Mode, Mode$TRVLCMIN >= 0)
Mode <- subset(Mode, Mode$TRPMILES >= 0)
Mode <- subset(Mode, Mode$HHFAMINC >= 0)
Mode <- subset(Mode, Mode$R_SEX >= 0)

Mode$R_SEX <- as.factor(Mode$R_SEX)
levels(Mode$R_SEX) <- c("0", "1") # 0 as male, 1 as female.

Mode$TRPTRANS <- as.factor(Mode$TRPTRANS)
levels(Mode$TRPTRANS) <- c("walk", "bicy", "pt", "pt", "taxi")

Mode.MNL <- Mode
Mode <- Mode[,c(-1,-2,-3)]

rm(walk)
rm(bicy)
rm(pt)
rm(taxi)

# plot(Mode, pch = 20, cex = 1)
```

```{r, echo=FALSE}
# how to split the data
# set.seed(123)
# 
# train <- sample(1:dim(Mode)[1], dim(Mode)[1]*0.7, rep=FALSE)
# test <- -train
# 
# training_data<- Mode[train, ]
# 
# testing_data= Mode[test, ]
```

### 1. Develop MNL ----
#### (1) MNL 1 - No variable is controled(All area; All kinds of trips) ----
```{r, echo=FALSE}
Mode$out <- relevel(Mode$TRPTRANS , ref = "walk")
library(nnet)
library(pscl)
library(mlogit)
#library(dplyr)

MNL1 <- multinom(out ~ TRVLCMIN + TRPMILES + GASPRICE + HHFAMINC + HHVEHCNT + HHSIZE + R_AGE + R_SEX + MSACAT + HBPPOPDN, data = Mode)
MNL1
a <- summary(MNL1)

  ### p-value
z <- summary(MNL1)$coefficients/summary(MNL1)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p

  ### Predict 
predict(MNL1, Mode)[1:30]
predict(MNL1, type = "prob")[1:30]
head(predict(MNL1, type = "prob"))


  ### Misclaasification error
cn1 <- table(predict(MNL1), Mode$out)
print(cn1) # off-diagonal is mis-classification

sum(diag(cn1))/sum(cn1) # overall classification correctness rate is 82.1%. However, walking contributes to the most of it. The actual prediction rate for other modes are relatively low. 

as.data.frame(diag(cn1))/apply(as.data.frame.matrix(cn1), 2, sum)
  

  ### r-suqare
round(pR2(MNL1),4)

```

#### (2) MNL 2 - Restricted to urban area and work trips ----
```{r,echo=FALSE}
# MNL 2 - Restricted to Urban and work
library(nnet)
library(pscl)
library(mlogit)
Mode.U.10 <- subset(Mode, as.character(Mode$HBHUR) == "U" & Mode$WHYTRP1S == 10)
table(Mode.U.10$TRPTRANS)
levels(Mode.U.10)


MNL2 <- multinom(TRPTRANS ~ TRVLCMIN + TRPMILES + GASPRICE + HHFAMINC + HHVEHCNT + HHSIZE + R_AGE + R_SEX + MSACAT + HBPPOPDN, data = Mode.U.10)
summary(MNL2)

  ### two-tailed z-test
z <- summary(MNL2)$coefficients/summary(MNL2)$standard.errors
p<- (1 - pnorm(abs(z), 0, 1)) * 2
p

  ### odd ratio
oddsML <- exp(coef(MNL2))
print(oddsML, digits =2)


round(pR2(MNL2),4)

LL0 <- -2561*log(3)
LL0


adr2 <- 1-(logLik(MNL2)-9)/-2472.1724
adr2

### chi2
-2*(LL0-logLik(MNL2))
qchisq(0.95,df=9)
MNL1.1 <- multinom(TRPTRANS ~ 1, data = Mode.U.10)
lrtest(MNL2, MNL1.1)



  ### Predict 
predict(MNL2, Mode.U.10)[1:30]
predict(MNL2, type = "prob")[1:30]
head(predict(MNL2, type = "prob"))

  ### Misclaasification error
cn2 <- table(predict(MNL2), Mode.U.10$TRPTRANS)
print(cn2) # off-diagonal is mis-classification

sum(diag(cn2))/sum(cn2) # overall classification errors are similar

as.data.frame(diag(cn2))/apply(as.data.frame.matrix(cn2), 2, sum)





  ### confusion matrix
caret::confusionMatrix(as.factor(predict(MNL2)), as.factor(Mode.U.10$TRPTRANS))


# hmftest(MNL1)

Myvars.RF <- c("TRPTRANS", "WHYTRP1S", "HBHUR", "TRVLCMIN", "TRPMILES", "GASPRICE", "HHFAMINC", "HHVEHCNT", "HHSIZE", "R_AGE", "R_SEX", "MSACAT", "HBPPOPDN")
Mode.RF <- Mode[Myvars.RF]
Mode.RF <- subset(Mode.RF, as.character(Mode.RF$HBHUR) == "U" & Mode.RF$WHYTRP1S == 10)
Mode.RF <- Mode.RF[,c(-2, -3)]


Train <- sample(1:dim(Mode.RF)[1], dim(Mode.RF)[1]*0.7, rep=FALSE)
Test <- -Train

training_data<- Mode.RF[Train, ]
testing_data= Mode.RF[Test, ]

TrainY=training_data[[1]]  
TrainX=training_data[,-1]

TestY=testing_data[[1]]
TestX=testing_data[,-1]


MNL3 <- multinom(TrainY ~ TRVLCMIN + TRPMILES + GASPRICE + HHFAMINC + HHVEHCNT + HHSIZE + R_AGE + R_SEX + MSACAT + HBPPOPDN, data = TrainX)



  ### two-tailed z-test
z <- summary(MNL3)$coefficients/summary(MNL3)$standard.errors
p<- (1 - pnorm(abs(z), 0, 1)) * 2
p

  ### odd ratio
oddsML <- exp(coef(MNL3))
print(oddsML, digits =2)


round(pR2(MNL3),4)

LL0 <- -nrow(TrainX)*log(3)
LL0


adr2 <- 1-(logLik(MNL3)-9)/-2472.1724
adr2

### chi2
-2*(LL0-logLik(MNL3))
qchisq(0.95,df=9)
MNL1.1 <- multinom(TrainY ~ 1, data = TrainX)
lrtest(MNL3, MNL1.1)


caret::confusionMatrix(as.factor(predict(MNL3,TestX)), as.factor(TestY))


```
# A MNL model tries to reduce r2
```{r}

MNL4 <- multinom(TrainY ~ TRVLCMIN + GASPRICE + HHFAMINC + HHVEHCNT + HHSIZE + R_AGE + R_SEX + MSACAT + HBPPOPDN , data = TrainX)

z <- summary(MNL4)$coefficients/summary(MNL4)$standard.errors
p<- (1 - pnorm(abs(z), 0, 1)) * 2
p

  ### odd ratio
oddsML <- exp(coef(MNL4))
print(oddsML, digits =2)


round(pR2(MNL4),4)

LL0 <- -nrow(TrainX)*log(3)
LL0


adr2 <- 1-(logLik(MNL4)-8)/-1749.7068
adr2

### chi2
-2*(LL0-logLik(MNL4))
qchisq(0.95,df=9)
MNL1.1 <- multinom(TrainY ~ 1, data = TrainX)
lrtest(MNL4, MNL1.1)

caret::confusionMatrix(as.factor(predict(MNL4,TestX)), as.factor(TestY))


```

### 2. Develop unsupervised learning ----
#### (1) Cluster analysis ----
```{r, echo=FALSE}
Myvars.Cst <- c("TRPTRANS", "WHYTRP1S", "HBHUR", "TRVLCMIN", "TRPMILES", "GASPRICE", "HHFAMINC", "HHVEHCNT", "HHSIZE", "R_AGE", "MSACAT", "HBPPOPDN")
Mode.Cst <- Mode[Myvars.Cst]



Mode.Cst <- Mode.Cst[,c(-2, -3)]

Mode.UrbWrk.Cst <- Mode[Myvars.Cst]
Mode.UrbWrk.Cst <- subset(Mode.UrbWrk.Cst, as.character(Mode$HBHUR) == "U" & Mode$WHYTRP1S == 10)
Mode.UrbWrk.Cst <- Mode.UrbWrk.Cst[,c(-2,-3)]

#plot(Mode.UrbWrk.Cst)                                     # Output 0

# Normalize ----
m <- Mode.UrbWrk.Cst[,-1] # without actual mode
mean <- apply(m, 2, mean)
sd <- apply(m, 2, sd)
Mode.UrbWrk.Cst.Scaled <- scale(m, mean, sd)

# K-MEANS CLUSTERING-----
## CLUSTERING
fitK <- kmeans(Mode.UrbWrk.Cst.Scaled, 4)
fitK
fitK$centers                                                     # Output 1
fitK$betweenss/fitK$tot.withinss

plot(Mode.UrbWrk.Cst, col = fitK$cluster)                  # Output 2


## CHOOSE K-----
k <- list()
for (i in 1:10){
  k[[i]] <- kmeans(Mode.UrbWrk.Cst.Scaled, i)
}

betweenss_totss <- list()
for (i in 1:10){
  betweenss_totss[[i]] <- k[[i]]$betweenss/k[[i]]$totss
}

plot(1:10, betweenss_totss,type = "b", ylab = "Between SS/ Total SS", xlab = "Cluster(k)")#look where the shoulder is   # Output 3

Mode.UrbWrk.Cst$Out <- fitK$cluster


as.data.frame(table(Mode.UrbWrk.Cst$TRPTRANS))                                                             # Output 4
as.data.frame(table(subset(Mode.UrbWrk.Cst, as.character(Mode.UrbWrk.Cst$TRPTRANS)=="taxi")$Out))          # Output 5
```



#### (2) PCA + Cluster ----
```{r, echo=FALSE}

pca.x <- data.frame(Mode.UrbWrk.Cst.Scaled)

library(psych)
library(SensoMineR)
fa.parallel(pca.x, fa = 'pc', n.iter = 100, show.legend = FALSE) #Scree plot                           # Output 6

#da.panelellipses=panellipse(pca.x, col.p=2, col.j=1, firstvar=3)

pca.fit <- principal(pca.x, nfactors = 4, rotate = "none")
(pca.fit)

pca.X <- as.data.frame(pca.fit$scores)
pca.X


pca.full <- data.frame(Mode.UrbWrk.Cst$TRPTRANS, pca.X)
plot(pca.full)                                                                                        # Output 7

# k-means----
fit.pca.k <- kmeans(pca.X, 4)
fit.pca.k
fit.pca.k$centers                                                     # Output 1
fit.pca.k$betweenss/fit.pca.k$tot.withinss


#table(Mode_C_U_W$TRPTRANS)
plot(pca.full, col = fit.pca.k$cluster)                                                               # Output 8
# plot(pca.full, col = pca.full$Mode.UrbWrk.Cst.TRPTRANS)
pca.full$out <- fit.pca.k$cluster


for (i in 1:10){
  k[[i]] <- kmeans(pca.X, i)
}

betweenss_totss <- list()
for (i in 1:10){
  betweenss_totss[[i]] <- k[[i]]$betweenss/k[[i]]$totss
}

plot(1:10, betweenss_totss,type = "b", ylab = "Between SS/ Total SS", xlab = "Cluster(k)")#look where the shoulder is   # Output 9    

as.data.frame(table(Mode.UrbWrk.Cst$TRPTRANS))                                                     # Output 10
as.data.frame(table(subset(pca.full, as.character(pca.full$Mode.UrbWrk.Cst.TRPTRANS)=="taxi")$out))       # Output 11

```
### 3. Naive Bayers Classifier ----
```{r, echo=FALSE}
library(e1071)




Mode.UrbWrk.NB <- Mode.RF

Naive.Bayes.Model <- naiveBayes(TRPTRANS ~ . , data = Mode.UrbWrk.NB)
Naive.Bayes.Model

pred <- predict(Naive.Bayes.Model, Mode.UrbWrk.NB)
cn <- table(pred, Mode.UrbWrk.NB$TRPTRANS)
print(cn) # off-diagonal is mis-classification

sum(diag(cn))/sum(cn) # overall classification errors are similar

as.data.frame(diag(cn))/apply(as.data.frame.matrix(cn), 2, sum)

caret::confusionMatrix(as.factor(predict(Naive.Bayes.Model, Mode.UrbWrk.NB)), Mode.UrbWrk.NB$TRPTRANS)




Myvars.RF <- c("TRPTRANS", "WHYTRP1S", "HBHUR", "TRVLCMIN", "TRPMILES", "GASPRICE", "HHFAMINC", "HHVEHCNT", "HHSIZE", "R_AGE", "R_SEX", "MSACAT", "HBPPOPDN")
Mode.RF <- Mode[Myvars.RF]
Mode.RF <- subset(Mode.RF, as.character(Mode.RF$HBHUR) == "U" & Mode.RF$WHYTRP1S == 10)
Mode.RF <- Mode.RF[,c(-2, -3)]


Train <- sample(1:dim(Mode.RF)[1], dim(Mode.RF)[1]*0.7, rep=FALSE)
Test <- -Train

training_data<- Mode.RF[Train, ]
testing_data= Mode.RF[Test, ]

TrainY=training_data[[1]]  
TrainX=training_data[,-1]

TestY=testing_data[[1]]
TestX=testing_data[,-1]


Naive.Bayes.Model <- naiveBayes(TrainY ~ . , data = TrainX)

caret::confusionMatrix(as.factor(predict(Naive.Bayes.Model,TestX)), as.factor(TestY))

```

### 4. Random Forest ----  
```{r, echo=FALSE}
library(caret)
# library(e1071)

# Myvars.RF <- c("TRPTRANS", "WHYTRP1S", "HBHUR", "TRVLCMIN", "TRPMILES", "GASPRICE", "HHFAMINC", "HHVEHCNT", "HHSIZE", "R_AGE", "R_SEX")
# Mode.RF <- Mode[Myvars.RF]
# Mode.RF <- subset(Mode.RF, as.character(Mode.RF$HBHUR) == "U" & Mode.RF$WHYTRP1S == 10)
# Mode.RF <- Mode.RF[,c(-2, -3)]



# Mode.RF <- Mode.UrbWrk.Cst[, -9]

set.seed(123)

Train <- sample(1:dim(Mode.RF)[1], dim(Mode.RF)[1]*0.7, rep=FALSE)
Test <- -Train

training_data<- Mode.RF[Train, ]
testing_data= Mode.RF[Test, ]

TrainY=training_data[[1]]  
TrainX=training_data[,-1]

TestY=testing_data[[1]]
TestX=testing_data[,-1]

indx <- createFolds(TrainY, k = 5, returnTrain = TRUE)
ctrl <- trainControl(method = "repeatedcv", number = 5, repeats = 10, index = indx)
mtryGrid <- data.frame(mtry = floor(seq(2, ncol(TrainX), length = 10)))

# rfTune <- train(x = TrainX, y = TrainY, method = "rf", tuneGrid = mtryGrid, ntree = 1000, importance = TRUE, trControl = ctrl)
rfTune <- train(x = TrainX, y = TrainY, method = "rf", tuneGrid = mtryGrid, ntree = 1000, importance = TRUE, trControl = ctrl)
rfTune

# testResults<-data.frame(obs=TestY, predict(rfTune, TestX))
testResults <- table(obs=TestY, predict(rfTune, TestX))
testResults


sum(diag(testResults))/sum(testResults)
as.data.frame(diag(testResults))/apply(as.data.frame.matrix(testResults), 1, sum)


plot(rfTune)# Plot the tuning results

rfImp <- varImp(rfTune, scale = FALSE)  #variable importance
rfImp
plot(rfImp)


caret::confusionMatrix(as.factor(predict(rfTune, TestX)), TestY)


```

